---
title: "README.Rmd"
output: html_document
---

output:
  html_document: default
  pdf_document: default
---

Preguntas

a. Examine the portion of the script you see below. In your own words what does this section do and why does it work? 
vonset = Get starting point: 2, 2: the point when the vowel starts
voffset = Get end point: 2, 2: the points when the vowel ends
durationV = voffset - vonset: vowel duration is the final of the vowel less the onset of the vowel
per20 = vonset + (durationV * 0.20): is the 20% of the total vowel duration
per35 = vonset + (durationV * 0.35): is the 35% of the total vowel duration
per50 = vonset + (durationV * 0.50): is the 50% of the total vowel duration
per65 = vonset + (durationV * 0.65): is the 65% of the total vowel duration
per80 = vonset + (durationV * 0.80): is the 80% of the total vowel duration

b. In a few short sentences describe the general outline of the script, what the purpose is, and how it acheives this purpose (hint: focus on the section dividers and the comments)
The script has three sections: 
b.1 set up output file: praat look for the file in the data folder
b.2 set up the loop: praat look for the stim folder a
b.3 run the loop: 

c. In a few short sentences describe how the segmenting procedure you used this week differs from that used in pa_2. What are the advantages and disadvantages?



cargar base de datos 
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
df1 = read_csv("data/vowel_data.csv")

```


Manipulate the dataframe as necessary so that you can calculate average F1/F2 centroids and trajectory length (include SD).

 media y SD de las variables
 
```{r} 

df1 %>%
    summarise(media_f1 = mean(f1_cent), sd_f1 = sd(f1_cent), media_f2 = mean(f2_cent), sd_f2 = sd(f2_cent),  meadia_tl = mean(tl), sd_tl = sd(tl)) 


```


trajectory length as a function of vowel and language
  
```{r}

  ggplot(df1, aes(fill= language, y=tl, x=vowel)) + 
  geom_bar(position="dodge", stat="identity")
  
```


F1 as a function of vowel and language
```{r}
ggplot(df1, aes(fill= language, y=f1_cent, x=vowel)) + 
  geom_bar(position="dodge", stat="identity")
```


F2 as a function of vowel and language
```{r}
ggplot(df1, aes(fill= language, y=f2_cent, x=vowel)) + 
  geom_bar(position="dodge", stat="identity")
```
